\Huge \textbf{Species sympatry constrains brain size evolution in Primates}

### 2021-06-25 Article/Article.Rmd
Update analysis diversification (Bayesian) + finish writing

---
title: "Primate Cognition and the Red Queen Hypothesis"
# date: "`r format(Sys.time(), '%B %d, %Y')`"

#I follow: https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown for author display
author: 
  - name: Benjamin Robira
    email: benjamin.robira@normalesup.org
    institute: [CEFE, MH]
    correspondence: true
  - name: Benoit Perez-Lamarque
    institute: [ENS, MNHN]
institute:
  - CEFE: Centre d’Ecologie Fonctionnelle et Evolutive, Université de Montpellier & CNRS, Montpellier, France.
  - MH: Anthropologie et Ethnobiologie, Centre National de la Recherche Scientifique/Muséum National d'Histoire Naturelle, University Paris Diderot, Sorbonne Paris Cité, Musée de l'Homme, Paris, France
  - ENS: Institut de Biologie de l’Ecole Normale Supérieure (IBENS), Ecole Normale Supérieure, CNRS, INSERM, Université PSL, Paris, France
  - MNHN: Institut de Systématique, Evolution, Biodiversité (ISYEB), Muséum National d’Histoire Naturelle, CNRS, Sorbonne Université, EPHE, Université des Antilles, Paris, France

@@

bibliography: bibliographyarticlepackage.bib
csl: science.csl
#biblio-style: apa
always_allow_html: true

@@

header-includes:
  - \usepackage[nolists, nomarkers,tablesfirst]{endfloat} # For figures and tables at end
  - \usepackage{lineno} # For line numbering
  - \linenumbers # For line numbering
  - \usepackage{setspace}\doublespacing
  - \setlength{\parskip}{0em} #to remove line gaps
  - \DeclareUnicodeCharacter{2212}{-}
  - \usepackage{caption}
  - \captionsetup[figure]{font=small}
---

@@

#Phylogeny
library(phytools)
library(ape)

@@

#Create citation fusion between articles and package (based on toolbox function)
citeR(
bibliographyArticle="C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Article/bibliographyarticle.bib",
bibliographyOutput="C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Article/bibliographyarticlepackage.bib",
rgeos,
geosphere,
phytools,
geiger,
RPANDA,
caper,
neurobase,
misc3d,
phylolm,
nlme,
pgls
)

@@

@@

```{r,  warning = FALSE, message = FALSE}
#For the moment because this was not run: to delete after
repetitionTrees=50
repetitionModels=30
repetitionBootstrap=200

repetition=7#length(frugivoryThresholdVector)*length(folivoryThresholdVector)*length(geographicThresholdVector)*randomSampling - 1
checkSampleFruit <- rep(NA, times=repetition)
checkSampleLeaf <- rep(NA, times=repetition)
checkSampleRange <- rep(NA, times=repetition)
checkSampleBrain <-  rep(NA, times=repetition)
checkSampleEQ <-  rep(NA, times=repetition)
checkSampleNeocortex <-  rep(NA, times=repetition)
checkSampleHippocampus <- rep(NA, times=repetition)
checkSampleCerebellum <- rep(NA, times=repetition)
checkSampleStriatum <- rep(NA, times=repetition)
checkSampleMOB <- rep(NA, times=repetition)

for(i in 0:repetition){

  summaryDataOfInterest <- read_delim(paste("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/OutputEvolModel/Data",i,".txt", sep=""), "\t", escape_double = FALSE, trim_ws = TRUE)

  checkSampleFruit[i] <- length(summaryDataOfInterest$DietaryGuild[summaryDataOfInterest$DietaryGuild=="Fruit"])
  checkSampleLeaf[i]  <- length(summaryDataOfInterest$DietaryGuild[summaryDataOfInterest$DietaryGuild=="Leaf"])
  checkSampleRange[i]  <- length(summaryDataOfInterest$geographicCode[!is.na(summaryDataOfInterest$geographicCode)])
  checkSampleBrain[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&summaryDataOfInterest$DietaryGuild=="Fruit",])
  checkSampleEQ[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$Bodymass)&summaryDataOfInterest$DietaryGuild=="Fruit",])
  checkSampleNeocortex[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$Neocortex)&summaryDataOfInterest$DietaryGuild=="Fruit",]) 
  checkSampleHippocampus[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$Hippocampus)&summaryDataOfInterest$DietaryGuild=="Fruit",])
  checkSampleCerebellum[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$Cerebellum)&summaryDataOfInterest$DietaryGuild=="Fruit",])
  checkSampleStriatum[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$Striatum)&summaryDataOfInterest$DietaryGuild=="Fruit",])
  checkSampleMOB[i]  <- nrow(summaryDataOfInterest[!is.na(summaryDataOfInterest$Brain)&!is.na(summaryDataOfInterest$MOB)&summaryDataOfInterest$DietaryGuild=="Fruit",])

@@

**Abstract:** The diversity in animal cognition raises the question of its underlying evolutionary drivers. Selection upon more advanced cognitive abilities can stem from interactions of individuals with conspecifics within the social unit (*Social Intelligence Hypothesis*), among generations (*Cultural Intelligence Hypothesis*), between social units (*Napoleonic Intelligence Hypothesis*), or with the rest of their environment (*Ecological Intelligence Hypothesis*). These hypotheses were limited to within-species scenarios. Yet, one species rarely occupies an area alone: space is a place shared between many species that can interact directly or indirectly. For species occupying a same ecological niche, interspecific competition is particularly expected in this case, but interactions could also represent additional cues and/or noise to cognitively process to find resource or escape predator(s) more efficiently. To test whether species co-occurrence shaped current patterns of cognition, we used Primates as a study example and retraced the evolutionary history of multiple species brain areas involved in foraging activities or not while considering competitive or non-competitive evolutionary scenarios. We found that the evolution of the relative size of areas involved in foraging-related information processing and/or retention is better described by models accounting for species co-occurrence within dietary guilds which lead to increased relative sizes of these areas. This suggests that cognition is subject to an arms race among species (*Red Queen Intelligence Hypothesis*) and shall explain why specific abilities particularly boomed for some lineages, as for instance with Primates.

**Keywords:** "Brain size - Species co-occurrence - Intelligence evolution - Cooperation - Competition"

```{r}
#My toolbox with several functions
source("T:/Saved_PhD/Empirical_analysis/Scripts&Functions/Functions/toolbox.R", local = knitr::knit_global())

words <- RmdWords("Article.Rmd")
```

To grasp the rationale of current hypotheses that aim to describe the evolution of animal cognition [@van2006some; @dunbar2017there], consider a primate individual foraging for food. Primates are pivotal species to study the evolution of cognition given the complexity of their socio-ecological environment and the inevitable implication for retracing human evolutionary history [@byrne2000evolution]. The chosen focal is not looking for any type of food but specific one, as for instance, most primate species often look for fruits. Finding fruits for a primate is not an easy task. In fact, fruits are the archetype of a hard-to-find resource likely to promote cognitive development because fruit trees are rare, dispersed, and do not produce constantly along the year nor between years, albeit their fruiting period remains nonetheless predictable [@janmaat2016spatio]. Additionally, they are energy-rich, allowing expansion of costly tissues such as brain tissues [@clutton1980primates]. Overall, primates thus show remarkable abilities to navigate precisely and target fruit trees likely to yield ripe resource despite they are out of their sensory range [@trapanese2019review], in part because of advanced spatio-temporal information retention [e.g. spatial knowledge: @normand2009forest; @robiraroute, temporal knowledge: @janmaat2006evidence; @janmaat2013chimpanzees; @janmaat2013tai].

|	When the resource is out of sight, moving at random could mean travelling unnecessary long distances to finally reach a potentially void tree. Hence, for the focal, being capable of processing immediate cues and clues to know where specifically to find fruit trees [@dall2005information; @grove2013evolution] and whether it is probable that these trees currently yield fruits [@dall2005information; @robirainreview; @janmaat2016spatio] would be a considerable advantage. Having appropriate cognitive machinery may avoid costly detours and time recursions to food patches adequately. This simple picture draws the basis of the *Ecological Intelligence Hypothesis* [@clutton1980primates; @milton1981distribution; @rosati2017foraging] which stipulates that cognition was selected as a way to buffer the spatio-temporal complexity of the environment (e.g. @van2012large).

|	In its quest for food, however, the focal might not be alone: for instance, primates often form social groups [@kappeler2002evolution] where group members spend considerable time together, have established bounds and constantly share information. Thus, they might altogether look for, and process, cues and clues leading to food sources. To do so, being able to process signals emitted by other maneuvers would be an advantage, as well as reading through their mind [@devaine2017reading] in order to plan ahead for collective actions or even outsmart them [@byrne1994machiavellian]. This therefore means plenty of additional stimuli to process, and constitutes the socle of the *Social Intelligence Hypothesis* [@dunbar2017there].

|   In the focal group, however, all individuals are not equally knowledgeable: perhaps because some have more experience than others do, such as the oldest would be more experienced than the youths. It is possible that these young individuals learn from the elderly ones [@laland2004social]. For instance, they may learn how to process a specific tool to access a hidden comestible part as for nuts [@boesch1994nut]. In other words, there is vertical transmission that passes by over generations and knowledge thereby accumulates: this is the *Cultural Intelligence Hypothesis* [@whiten2007evolution, reader2002social], also known as the *Vygotskian Intelligence Hypothesis* in humans [@herrmann2007humans; @tomasello2009cultural; @wilson1991molecular; @van2011social].

|   In this overly simplistic picture, we moved from a unique individual to a group of individuals and then to multiple groups. Never was it question of multiple species: all hypothesized individuals belonged to a singular common species. Yet, space is a place shared between a plethora of species, some of them occupying a same ecological niche. As much as conspecifics could be competitors, or direct or indirect cooperators, so could be individuals from another species with similar ecological preferences, for instance with regards to diet. As such, co-occurrence of species from the same guild might contribute into shaping animal cognition: the more species, the more selective pressure for advanced cognition. In this line, because the cognitive abilities would rhyme with ecological success (REF), thus inducing higher population size which ultimately would spur speciations and reduce extinction risk, these cognitively advanced species are expected to "boom", what would be translated into increased species diversification rates. In this study, we therefore propose to test for the existence of such an arms race [@van1973new] in cognitive abilities among species - approximated by the size of the whole or specific regions of the brain - that we naturally term the *Red Queen Intelligence Hypothesis*, building on frugivorous primates as a study example, and assess how this might have affected the evolutionary success of the Primate lineage.

>>>>

---
title: \Huge \textbf{Species sympatry constrains brain size evolution in Primates}
# date: "`r format(Sys.time(), '%B %d, %Y')`"

#I follow: https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown for author display
author: 
  - name: \textit{Benjamin Robira}
    email: benjamin.robira@normalesup.org
    institute: [CEFE, MH]
    correspondence: true
  - name: \textit{Benoit Perez-Lamarque}
    email: benoit.perez@ens.psl.eu
    institute: [ENS, MNHN]
    correspondence: true
institute:
  - CEFE: Centre d’Écologie Fonctionnelle et Évolutive, Université de Montpellier & CNRS, Montpellier, France.
  - MH: Anthropologie et Ethnobiologie, Centre National de la Recherche Scientifique/Muséum National d'Histoire Naturelle, University Paris Diderot, Sorbonne Paris Cité, Musée de l'Homme, Paris, France
  - ENS: Institut de Biologie de l’École Normale Supérieure (IBENS), École Normale Supérieure, CNRS, INSERM, Université PSL, Paris, France
  - MNHN: Institut de Systématique, Évolution, Biodiversité (ISYEB), Muséum National d’Histoire Naturelle, CNRS, Sorbonne Université, EPHE, Université des Antilles, Paris, France

@@

bibliography: bibliographyarticlepackage.bib
csl: science.csl
#biblio-style: apa
always_allow_html: true
link-citations: yes

@@

header-includes:
# - \usepackage[nolists, nomarkers,tablesfirst]{endfloat} # For figures and tables at end
  - \usepackage{lineno} # For line numbering
  - \linenumbers # For line numbering
  - \usepackage{setspace}\doublespacing
  - \setlength{\parskip}{0em} #to remove line gaps
  - \DeclareUnicodeCharacter{2212}{-}
  - \usepackage{caption}
  - \captionsetup[figure]{font=small}
  - \newcommand{\beginsupplement}{
      \setcounter{table}{0}  
      \renewcommand{\thetable}{S\arabic{table}} 
      \setcounter{figure}{0} 
      \renewcommand{\thefigure}{S\arabic{figure}}
    }
---

@@

#Phylogeny
library(phytools)
library(ape)
library(phylolm)

#Segmentation
library(strucchange)

@@

#Create citation fusion between articles and package (based on toolbox function)
citeR(
bibliographyArticle="C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Article/bibliographyarticle.bib",
bibliographyOutput="C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/Article/bibliographyarticlepackage.bib",
rgeos,
geosphere,
phytools,
geiger,
RPANDA,
caper,
neurobase,
misc3d,
phylolm,
nlme,
MCMCglmm,
coda,
strucchange
)

@@

load("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/REnvironments/PGLSdiversification_withautocorr.RData")
load("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates/REnvironments/PGLSdirectionSelection.RData")

@@

**Abstract:** The diversity in animal cognition raises the question of its underlying evolutionary drivers. Selection upon more advanced cognitive abilities can stem from interactions of individuals with conspecifics within the social unit (*Social Intelligence Hypothesis*), among generations (*Cultural Intelligence Hypothesis*), between social units (*Napoleonic Intelligence Hypothesis*), or with the rest of their environment (*Ecological Intelligence Hypothesis*). These hypotheses were limited to within-species scenarios. Yet, one species rarely occupies an area alone: Space is a place shared between many species that can interact directly or indirectly. For species occupying a same dietary niche, interspecific competition could induce a cognitive evolutionary arms race, so as to outperform competitors. Furthermore, all species act upon the spatio-temporal distribution of resources, hence contribute to increasing the uncertainty for a third species to infer resource location and availability. This would ultimately shape the cognitive machinery involved to allow sufficient harvesting. In addition, sympatric species can leave traces of passage and of resource presence that can be cognitively processed by a third species to find it more efficiently. As such, to test whether species co-occurrence shaped current patterns of cognition, we used primates as a study example and retraced the evolutionary history of multiple species brain areas involved in foraging activities or not while considering competitive or non-competitive evolutionary scenarios. We found that the evolution of the relative size of areas involved in foraging-related information processing and/or retention, as well as in areas related to processing social information, are better described by models accounting for species co-occurrence within dietary guilds. More precisely for these brain areas, species co-occurrence was associated to a decrease, and never to an increase, of their relative size. Coherently with the observed wide variability in sympatry rate and intensity, the degree of encephalisation was unrelated to the evolutionary success of a lineage (i.e. its diversification rate). Overall, this comparative study suggests that species co-occurrence stands as a brake to positive selection towards larger cognitive abilities, yet leaves open the question of the underlying ecological mechanisms at play.

<!-- subject to an arms race among species (*Red Queen Intelligence Hypothesis*).  -->

**Keywords:** "Brain size - Cognition - Cooperation - Competition - Intelligence evolution - Primates - Species co-occurrence"

<!-- Reviewer possibles: Decasien, Powell, Barton, Kamilar, Harmon, Drury -->

```{r wordCount}
words <- RmdWords("Article.Rmd")
```

# Introduction

On the road to brain size evolution, generally considered as an equivalent of cognition evolution, mysteries are plenty [@dunbar2017there]. It remains puzzling why humans have a brain that is that large, relatively to body size, compared to other animals, or why primate brain architecture, in general, is much more complex than observed in other taxa. Overall, the evolution of the brain is the consequence of constraints [e.g. energetic limitations, @navarrete2011energetics], but also of socio-ecological drivers promoting cognitive abilities [@gonzalez2018inference].

|   To grasp the rationale of current hypotheses that aim to describe the evolution of cognition [@van2006some; @dunbar2017there], consider a primate individual foraging for food. Primates are pivotal species to study the evolution of cognition given the complexity of their socio-ecological environment and the inevitable implication for retracing human evolutionary history [@byrne2000evolution]. The chosen focal is not looking for any type of food but specific one, as for instance, most primate species often look for fruits. Finding fruits for a primate is not an easy task. In fact, fruits are the archetype of a hard-to-find resource likely to promote cognitive development because fruit trees are rare, dispersed, and do not produce constantly along the year nor between years, albeit their fruiting period remains nonetheless predictable [@janmaat2016spatio]. Additionally, they are energy-rich, allowing expansion of costly tissues such as brain tissues [@clutton1980primates]. Overall, primates thus show remarkable abilities to navigate precisely and target fruit trees likely to yield ripe resource despite they are out of their sensory range [@trapanese2019review], in part because of advanced spatio-temporal information retention [e.g. spatial knowledge: @normand2009forest; @robiraroute, temporal knowledge: @janmaat2006evidence; @janmaat2013chimpanzees; @janmaat2013tai].

|   When the resource is out of sight, moving at random could mean travelling unnecessary long distances to finally reach a potentially void tree. Hence, for the focal, being capable of processing immediate environmental cues and clues to know where specifically to find fruit trees [@dall2005information; @grove2013evolution] and whether it is probable that these trees currently yield fruits [@dall2005information; @robirainreview; @janmaat2016spatio] would be a considerable advantage. Having an appropriate cognitive machinery may avoid costly detours and time recursions to food patches adequately. This simple picture draws the basis of the *Ecological Intelligence Hypothesis* [@clutton1980primates; @milton1981distribution; @rosati2017foraging] which stipulates that cognition was selected as a way to buffer the spatio-temporal complexity of the environment (e.g. @van2012large).

|   In its quest for food, however, the focal might not be alone: For instance, primates often form social groups [@kappeler2002evolution] in which group members spend considerable time together, have established bounds and constantly share information. Thus, they might altogether look for, and process, cues and clues leading to food sources. To do so, being able to process signals emitted by other maneuvers would be an advantage, as well as reading through their mind [@devaine2017reading] in order to plan ahead for collective actions or even outsmart them [@byrne1994machiavellian]. This therefore means plenty of additional information to process, and constitutes the socle of the *Social Intelligence Hypothesis* [@dunbar2017there].

|   In the focal group, however, all individuals are not equally knowledgeable: Perhaps because some have more experience than others, such as the elderly ones would be more experienced than the youths, with the latter thus learning from the former [@laland2004social]. For instance, young individuals may progressively learn how to process a specific tool to access a hidden comestible part as for nuts [@boesch1994nut] by observing and reproducing, or possibly being teached (learning ways detailed in @estienne2019acquisition). In other words, there is vertical transmission that passes by over generations and knowledge thereby accumulates: This is the *Cultural Intelligence Hypothesis* [@whiten2007evolution; @reader2002social], also known as the *Vygotskian Intelligence Hypothesis* in humans [@herrmann2007humans; @tomasello2009cultural; @wilson1991molecular; @van2011social].

|   Yet, in this overly simplistic picture, we moved from a unique individual to a group of individuals and then to multiple groups. Never was it question of multiple species. Yet, space is a place shared between a plethora of species, some of them occupying a same ecological niche. As much as conspecifics could be competitors, or direct or indirect cooperators, so could be individuals from another species with similar ecological preferences, for instance with regards to diet. As such, co-occurrence of species from the same guild might contribute into shaping animal cognition. On the one side, the presence of sympatric species with overlapping diet could contribute in reinforcing selective pressures for advanced cognition because (i, competition) species would compete for food access (i.e. Red Queen paradigm, [@van1973new]). Co-occurrence would increase the environmental complexity due to impoverishment of food and addition of noise to the spatio-temporal availability signal because of unforeseen depletion. (ii, cooperation/exploitation) sympatric species presence cues could also represent additional information to process to infer resource location and availability. Logically then, species with higher brain size would be the most evolutionary successful, thus “booming” and intensively diversifying as for hominins [@melchionna2020macroevolutionary]. On the other side, (i) the increase in environmental complexity could be such that advanced cognition is no longer adaptive (see for instance [@grove2013evolution] and [@robirainreview] for the limit of the adaptiveness of spatial and temporal cognition respectively), or (ii) the additional cues provided by other species presence would not add, but replace, and perhaps be more easily interpreted than, environmental cues of food availability. In this latter case, selective pressure on cognitive abilities would be relaxed, inducing a decrease in brain size in sympatric species compared to lonely species. Following this rationale, species living in co-occurrence or not with other species would not face the same evolutionary fate despite initially similar environmental conditions. Thus, the evolutionary success of specific lineages would end up unrelated to brain size. In this study, we therefore aimed to test how species co-occurrence contributed to shaping the evolutionary history of the encephalisation of the whole, or part of, the brain, and whether this induced a "boom" or a "brake" to cognitive abilities and associated evolutionary success by focusing on frugivorous primates as a study example.

```{r sampleCalculation,  warning = FALSE, message = FALSE}
###Set working directory
setwd("C:/Users/robira/Documents/PhD/Meta_analysis/Meta_analysis_cognition_primates")

repetition=2*2*2*10#length(frugivoryThresholdVector)*length(folivoryThresholdVector)*length(geographicThresholdVector)*randomSampling

checkSampleFruit <- rep(NA, times=repetition)
checkSampleLeaf <- rep(NA, times=repetition)
checkSampleRange <- rep(NA, times=repetition)
checkSampleBrain <-  rep(NA, times=repetition)
checkSampleEQ <-  rep(NA, times=repetition)
checkSampleNeocortex <-  rep(NA, times=repetition)
checkSampleHippocampus <- rep(NA, times=repetition)
checkSampleCerebellum <- rep(NA, times=repetition)
checkSampleStriatum <- rep(NA, times=repetition)
checkSampleMOB <- rep(NA, times=repetition)
checkSampleRange <- rep(NA, times=repetition)

counter=0
for(a in 1:2){
  for(b in 1:2){
    for(c in 1:2){
      for(d in 1:10){
        counter=counter+1
        tryCatch(
        {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleFruit",a,"_",b,"_",c,"_",d, ".txt", sep=""))
         checkSampleFruit[counter] <- toAdd[1]
        }, error=function(e){
          #Do nothing
        }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleLeaf",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleLeaf[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleRange",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleRange[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleBrain",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleBrain[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleEQ",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleEQ[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleNeocortex",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleNeocortex[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleHippocampus",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleHippocampus[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleCerebellum",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleCerebellum[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleStriatum",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleStriatum[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
        tryCatch(
          {toAdd <- read.delim(paste("Processed_data/Sample_size/checkSampleMOB",a,"_",b,"_",c,"_",d, ".txt", sep=""))
           checkSampleMOB[counter] <- toAdd[1]
          }, error=function(e){
            #Do nothing
          }
        )
      }
    }
  }
}

checkSampleFruit <- unlist(checkSampleFruit)
checkSampleLeaf <- unlist(checkSampleLeaf)
checkSampleRange <- unlist(checkSampleRange)
checkSampleBrain <-  unlist(checkSampleBrain)
checkSampleEQ <-  unlist(checkSampleEQ)
checkSampleNeocortex <-  unlist(checkSampleNeocortex)
checkSampleHippocampus <- unlist(checkSampleHippocampus)
checkSampleCerebellum <- unlist(checkSampleCerebellum)
checkSampleStriatum <- unlist(checkSampleStriatum)
checkSampleMOB <- unlist(checkSampleMOB)

@@

# Results

|   Recent tools have been developed to infer the effect of species interactions on trait evolution, either by modelling trait divergence in co-occurring species from a same guild (e.g. dietary guild; Matching Competition: MC models) or considering that the evolutionary rate depends on the number of lineages within the guild (density dependence; linear: DD$_{lin}$ or exponential: DD$_{exp}$; @drury2016estimating). After reconstructing primate biogeography history when considering `r length(areaName)` biogeographic areas following [@kamilar2009environmental] based on `r minRange` primate species (@matzke2013probabilistic; @matzke2016stochastic; Figure \@ref(fig:figmap)) as well as primate diet evolution based on `r minFruit + minLeaf` to `r maxFruit + maxLeaf` species (discrete trait: frugivory vs. folivory; @bollback2006simmap. The classification varied depending on how frugivory/folivory was assessed, see [Dietary guild]), we calculated the likelihoods of models considering the role of species interactions (including competitive scenarios) in the evolution of either the whole brain (using the encephalic quotient, EQ, as a proxy for `r minEQ` to `r maxEQ` frugivorous), or the relative size of specific brain areas associated with foraging-related information perception, processing or retention (Figure. \@ref(fig:figbrain); comprising `r minAllAreas` to `r max(maxNeocortex,maxHippocampus,maxCerebellum,maxStriatum,maxMOB)` frugivorous species). The use of specific region size relatively to the body mass (see @deaner2000comparative for further consideration of scaling methods), and not raw size, rather depicts the evolutionary evolution of cognitive abilities in terms of allocation rather than abilities per se (although it is vividly discussed whether raw measures are anything informative on "abilities" too @logan2018beyond). We also estimated the likelihoods of simpler models assuming no effect of species interactions, like the Brownian Motion (BM), the Ornstein-Uhlenbeck process (OU) assuming that traits are constrained around on optimal value (e.g. stabilizing selection; see @blomberg2020beyond for a review on these approaches) or the Early-Burst model (EB, @blomberg2003testing), this latter allowing to check for a time-dependence of the evolutionary rate, hence emphasizing that, if any, the density effect is not an artefact due to time dependence. Support for each model was evaluated using an information-theoretic framework [@burnham2002model] based on the weight of Akaike Information Criterion corrected for small samples (AICc) when considering all six models (MC, DD$_{lin}$, DD$_{exp}$, BM, OU, EB, see [Models of trait evolution: does interspecific interactions shape brain size evolution?]). Non-competitive models were the most likely in describing the evolutionary history of the EQ, the Neocortex and the Cerebellum (Figure \@ref(fig:figbrain) and \@ref(fig:figresultsevolution)), two areas specifically involved in movement and/or immediate information processing [@wiltgen2004new; @koziol2014consensus; @sokolov2017cerebellum] but also in memory consolidation for the Neocortex [@wiltgen2004new]. By contrast, competitive models were most supported in areas involved in sensory abilities (the main olfactory bulb, MOB), short-term working memory and long-term spatio-temporal information retention (Hippocampus, @burgess2002human), and the Striatum an area that is involved in information processing during social interaction (i.e. social reward assessment; @izuma2008processing) (Figure \@ref(fig:figbrain) and \@ref(fig:figresultsevolution)). When density-dependent models were the best fit, the rate (*r*, Figure \@ref(fig:figresultsevolution)) suggested an acceleration of the evolutionary tempo together with increased lineage diversity for the Hippocampus and Striatum, but a slow down for the MOB.

```{r samplecooccurrence}

minNspeciesCoocc <- round(min(dataRangePrimate$Number_species_cooccurrence, na.rm=TRUE), digit=2)
maxNspeciesCoocc <- round(max(dataRangePrimate$Number_species_cooccurrence, na.rm=TRUE), digit=2)
meanNspeciesCoocc <- round(mean(dataRangePrimate$Number_species_cooccurrence, na.rm=TRUE), digit=2)
seNspeciesCoocc <- round(sd(dataRangePrimate$Number_species_cooccurrence, na.rm=TRUE)/sqrt(length(dataRangePrimate$Number_species_cooccurrence[!is.na(dataRangePrimate$Number_species_cooccurrence)])), digit=2)

whichSpeciesMinNCoocc <- paste(sort(dataRangePrimate$Species[which(round(dataRangePrimate$Number_species_cooccurrence, digit=2)==minNspeciesCoocc)]), collapse=", ")
whichSpeciesMaxNCoocc <- paste(sort(dataRangePrimate$Species[which(round(dataRangePrimate$Number_species_cooccurrence, digit=2)==maxNspeciesCoocc)]), collapse=", ")
whichSpeciesMinNCoocc <- gsub("_", " ", whichSpeciesMinNCoocc)
whichSpeciesMaxNCoocc <- gsub("_", " ", whichSpeciesMaxNCoocc)

minOverlapSpecies <- round(min(dataRangePrimate$Overlap_average, na.rm=TRUE), digit=2)
maxOverlapSpecies <- round(max(dataRangePrimate$Overlap_average, na.rm=TRUE), digit=2)
meanOverlap <- round(mean(dataRangePrimate$Overlap_average, na.rm=TRUE), digit=2)
seOverlap <- round(sd(dataRangePrimate$Overlap_average, na.rm=TRUE)/sqrt(length(dataRangePrimate$Overlap_average[!is.na(dataRangePrimate$Overlap_average)])), digit=2)

whichSpeciesMinOverlap <- paste(dataRangePrimate$Species[which(round(dataRangePrimate$Overlap_average, digit=2)==minOverlapSpecies)], collapse=", ")
whichSpeciesMaxOverlap <- paste(dataRangePrimate$Species[which(round(dataRangePrimate$Overlap_average, digit=2)==maxOverlapSpecies)], collapse=", ")
whichSpeciesMinOverlap <- gsub("_", " ", whichSpeciesMinOverlap)
whichSpeciesMaxOverlap <- gsub("_", " ", whichSpeciesMaxOverlap)

```{r estimateRegGradient, error=TRUE, include=TRUE}

#Hippocampus
model <- get(paste("modelBrain", traitName[3], sep="_"))

estimateGradientHippocampus <- textEstOneVar(summary(model)$coefficients[2,c(1,4,5)])
testGradientHippocampus <- textTestOneVar(summary(model)$coefficients[2,c(3,6)], statistics="")

#Striatum
model <- get(paste("modelBrain", traitName[6], sep="_"))

estimateGradientStriatum <- textEstOneVar(summary(model)$coefficients[2,c(1,4,5)])
testGradientStriatum <- textTestOneVar(summary(model)$coefficients[2,c(3,6)], statistics="")

#MOB
model <- get(paste("modelBrain", traitName[7], sep="_"))

estimateGradientMOB <- textEstOneVar(summary(model)$coefficients[2,c(1,4,5)])
testGradientMOB <- textTestOneVar(summary(model)$coefficients[2,c(3,6)], statistics="")
```

|   Next, to understand the directionality of the selection gradient shaped by co-occurrence (i.e. selection for “bigger” or “smaller” brain if the more species), we fitted phylogenetic regressions (see [Phylogenetic regressions] a)). For these linear regressions, the predicted variable was the relative brain size values of the different areas. We considered the average surface of the frugivorous species range that was overlapped by other sympatric frugivorous species, as well as the number of such sympatric frugivorous species across their entire range as covariates. On average ($\pm$ SE), the considered primate species had `r meanNspeciesCoocc` in their distribution range ($\pm$ `r seNspeciesCoocc`). That ranged from `r minNspeciesCoocc` other species (*`r whichSpeciesMinNCoocc`*), to `r maxNspeciesCoocc` species (*`r whichSpeciesMaxNCoocc`*). On average ($\pm$ SE), the considered primate species had `r meanOverlap*100`% of their range overlapping with other species ($\pm$ `r seOverlap*100`). That ranged from `r minOverlapSpecies*100`% of overlap (*`r whichSpeciesMinOverlap`*), to `r maxOverlapSpecies*100`% of overlap (*`r whichSpeciesMaxOverlap`*). The number of sympatric species never influenced significantly the relative size of the brain or other specific areas (Table \@ref(tab:tableRegGradient)). The percentage of range shared on average with other species was coherently assessed as significantly correlated, or as a trend, with the relative size of areas which evolutionary history was better described with competitive models: the Hippocampus, the MOB and the Striatum (Hippocampus: $t$`r testGradientHippocampus`;  MOB: $t$`r testGradientMOB`; Striatum: $t$`r testGradientStriatum`). The correlations were all negative (Hippocampus: `r estimateGradientHippocampus`;  MOB: `r estimateGradientMOB`; Striatum: `r estimateGradientStriatum`), which means that higher overlap rhymes with lower relative size, insensitive to data and the phylogeny variability (Table \@ref(tab:tabledfsensitivity)). Thus, it suggests that sympatric species are subject to less stringent selection on advanced cognitive abilities.

```{r estimateRegDiversification, error=TRUE, include=TRUE}
